### 2018-01-18 Chapters/Chapter01.tex
Update blockchain

>>>>

\chapter{Analysis of Bitcoin}

\section{Hash function}

A hash function $h: \{0, 1\}^\infty \rightarrow \{0, 1\}^{n}$ has the following properties:

\begin{enumerate}
	\item $x = y \Rightarrow h(x) = h(y)$
	\item $h(x) \sim \mathcal{U}(0, 2^{n}-1)$, where $\mathcal{U}$ is the uniform distribution, i.e., $\forall a \in [0, 2^{n}-1], \mathbf{P}(h(x)=a) = \frac{1}{2^{n}}$
\end{enumerate}

In other words, when two inputs are the same, they have the same output. But, when the inputs are different, their outputs are uniformly distributed. Clearly, the hash functions are surjective but not injective. They are not injective because the image of $h$ has only $2^n$ elements and the domain has infinite elements. When $x \ne y$ and $h(x) = h(y)$, we say that $x$ and $y$ are a collision. A hash function is considered to be safe when it is unknown how to find a collision of a given hash.

\section{Bitcoin mining}

Let $\mathbb{B}$ be the set of Bitcoin blocks and $h: \mathbb{B} \rightarrow \{0, 1\}^{256}$ be the Bitcoin hash function. The mining process consists of finding $x \in \mathbb{B}$ such as $h(x) < A$, where $A$ is a given number. The lesser the $A$, the harder to find a new block. In fact, $\mathbf{P}(h(x) < A) = \frac{A}{2^{256}}$.

Hence, in order to find a new block, one must try different inputs ($x_1, x_2, \dots, x_k$) until they find a solution, i.e., all attempts will fail ($h(x_i) \geq A$ for $i < k$) but the last ($h(x_k) < A$). The probability of finding a solution exactly in the $k^{th}$ attempt follows a geometric distribution. Let $X$ be the number of attempts until a success, then $\mathbf{P}(X = k) = (1-p)^{k-1} p$, where $p = \frac{A}{2^{256}}$. Also, we have $\mathbf{P}(X \leq k) = 1 - (1-p)^k$. The average number of attempts is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = \frac{1-p}{p^2}$.

%As one may fail many times until they find a solution, the probability of finding a solution in the $K$ trial follows a geometric distribution. Let $X$ be the number of trials until a success, then $\mathbf{P}(X = k) = (1-p)^{k-1} p$, where $p = \mathbf{P}(a < A)$, i.e., there would be $k-1$ failures until a success in the $k-th$ trial. The average number of trials is $\mathbf{E}(X) = 1/p$ and the variance is $\mathbf{V}(X) = (1-p)/p^2$.

In the Bitcoin's protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average. Suppose that the Bitcoin's network is able to calculate $H$ hashes per second -- $H$ is the total hash rate of the network. The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block. So, the rule of finding a new block every 10 minutes (600 seconds) -- on average -- leads to the following equation: $\mathbf{E}(T) = 600$. So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = 600 \Rightarrow p = \frac{1}{600H}$. Finally, $\mathbf{E}(X) = 600H$, $\mathbf{E}(T) = 600$, $\mathbf{V}(X) = (600H)^2 - 600H$, and $\mathbf{V}(T) = 600^2 - 600/H$.

The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{600H} \right)^{tH}$. But, as the Bitcoin's network hash rate is really big, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{600}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{600}$.

\begin{theorem}
	When $H \rightarrow +\infty$, the time between blocks follows an exponential distribution with parameter $\lambda = \frac{1}{600}$, i.e., $\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{600}}$.
\end{theorem}
\begin{proof}
\begin{align*}
	\mathbf{P}(T \leq t) &= 1 - (1-p)^{tH} \\
		     &= 1 - \left( 1 - \frac{1}{600H} \right)^{tH} \\
	\text{Replacing $u = 600H$,} \\
	\lim_{H \rightarrow +\infty} \mathbf{P}(T \leq t) &= \lim_{u \rightarrow +\infty} 1 - \left( 1 - \frac{1}{u} \right)^{\frac{tu}{600}} \\
				&= \lim_{u \rightarrow +\infty} 1 - \left[ \left( 1 - \frac{1}{u} \right)^{u} \right]^{\frac{t}{600}} \\
				&= 1 - \left( 1/e \right)^{\frac{t}{600}} \\
				&= 1 - e^{-\frac{t}{600}}
\end{align*}
\end{proof}

% A partir de qual valor de H é razoável dizer que T segue uma distribuição exponencial?
\begin{theorem}
	$x > M \Rightarrow |(1+1/x)^x - e| < e/M$.
\end{theorem}
\begin{proof}
	Let's use the classical inequality $\frac{x}{1+x} < \log(1+x) < x$ for $x > -1$. So, $\frac{1/x}{1+1/x} < \log(1+x) < 1/x$. Simplifying, $\frac{1/x}{1+1/x} = 1/(1+x)$.  Thus, $1/(1+x) < \log(1+1/x) < 1/x \Rightarrow x/(1+x) < x \log(1 + 1/x) < 1$.

	As $\log(1 + \frac{1}{M}) > 0$, $1 < 1 + \log(1 + \frac{1}{M})$.

	$x > M \Rightarrow 1/x < 1/M \Rightarrow 1+1/x < 1+1/M \Rightarrow 1/(1+1/x) > 1/(1+1/M) \Rightarrow x/(1+x) > M/(1+M)$.

	Again, $\log(1+x) < x \Rightarrow \log(1-1/M) < -1/M \Rightarrow 1 + \log(1-1/M) < (M-1)/M < M/(1+M)$, since $(x-1)/x < x/(x+1)$.

	Hence, $1 + \log(1-1/M) < M/(1+M) < x/(1+x) < x\log(1+1/x)$, and $x\log(1+1/x) < 1 < 1 + \log(1 + \frac{1}{M})$.

	Finally,

	\begin{align*}
		1 + \log(1-1/M) &< x\log(1+1/x) < 1 + \log(1 + \frac{1}{M}) \\
		e^{1 + \log(1-1/M)} &< e^{x\log(1+1/x)} < e^{1 + \log(1 + \frac{1}{M})} \\
		e \cdot e^{\log(1-1/M)} &< e^{\log((1+1/x)^x)} < e \cdot e^{\log(1 + \frac{1}{M})} \\
		e(1-1/M) &< (1+1/x)^x < e(1 + \frac{1}{M}) \\
		e-e/M &< (1+1/x)^x < e + e/M \\
		-e/M &< (1+1/x)^x - e < e/M \\
	\end{align*}
	Therefore, $\left| \left( 1+1/x \right)^x - e \right| < e/M$.
\end{proof}

We may consider $H$ big enough to say that $T$ follows an exponential distribution when $e/H < \epsilon$, where $\epsilon$ is the maximum approximation error. When $\epsilon = 10^{-6} \Rightarrow H > e \cdot 10^6$. So, when $H > 2.6 \text{Mh/s}$, our approximation is good enough.

Let's analyze the probability of finding a new block for a miner who has $\alpha$ percent of the network's total hash rate. Let $T_\alpha = \frac{X}{\alpha H}$ be the time required for the miner to find a new block. As $T_\alpha = \left( \frac{1}{\alpha} \right) T$, when $H \rightarrow +\infty$, $T_\alpha$ also follows an exponential with parameter $\lambda_\alpha = \frac{\alpha}{600}$. Hence, we confirm the intuition that the miner with $\alpha$ percent of the network's total hash power will find $\alpha$ percent of the blocks.

\begin{theorem}
	When the miner with $\alpha$ percent of the network's total hash rate is part of the mining network, $\mathbf{P}(\text{next block is from $T_\alpha$}) = \alpha.$
\end{theorem}
\begin{proof}
\begin{align*}
	\mathbf{P}(\text{next block is from $T_\alpha$}) &= P \left( T_\alpha = \min\{T_\alpha, T_{1-\alpha}\} \right) \\
		&= \frac{\lambda_\alpha}{\lambda_\alpha + \lambda_{1-\alpha}} \\
		&= \frac{\alpha/600}{\alpha/600 + (1-\alpha)/600} \\
		&= \frac{\alpha}{\alpha + 1 - \alpha} \\
		&= \alpha.
\end{align*}
\end{proof}

\begin{theorem}
	When one miner with $\alpha$ percent of the network's total hash rate multiplies their hash rate by $m$, the probability of finding the next block is multiplied by $\frac{m}{m \alpha + 1 - \alpha}$.
	\label{thm-miner-multiply}
\end{theorem}
\begin{proof}
	When miners increase their hash rate, they also increase the network's total hash rate. Let $H$ be the network's hash rate before the increase. Thus, the network's total hash rate after the increase is $H + (m-1) \alpha H = (1 - \alpha + m \alpha) H$. So,
\begin{align*}
	\mathbf{P}(\text{next block is from $T_{m \alpha}$}) &= P \left( T_{m \alpha} = \min\{T_{m \alpha}, T_{1-\alpha}\} \right) \\
		&= \frac{\lambda_{m \alpha}}{\lambda_{m \alpha} + \lambda_{1-\alpha}} \\
		&= \frac{m \alpha/600}{m \alpha/600 + (1-\alpha)/600} \\
		&= \frac{m \alpha}{m \alpha + 1-\alpha} \\
		&= \alpha \left( \frac{m}{m \alpha + 1 - \alpha} \right).
\end{align*}
\end{proof}

\begin{cor}
	If one miner has a really tiny percent of the network's total hash rate, then multiplying their hash rate by $m$ approximatelly multiplies their probability of finding the next block by $m$.
\end{cor}
\begin{proof}
	$$\lim_{\alpha \rightarrow 0} \mathbf{P}(\text{next block is from $T_{m \alpha}$}) = \lim_{\alpha \rightarrow 0} \frac{m}{m \alpha + 1 - \alpha} = m.$$
\end{proof}

That way, it is not exacly correct to say that when one doubles their hash rate, their probability will double as well. It is only true for small miners.

%Then, $\mathbf{P}(t_0 < T < t_1) = \mathbf{P}(t_0 < X/H < t_1) = \mathbf{P}(t_0 H < X < t_1 H) = [1 - (1-p)^{t_1 H}] - [1 - (1-p)^{t_0 H}] = (1-p)^{t_0 H} - (1-p)^{t_1 H}$.

The symmetrical confidence interval with level $\alpha$ would be $[t_0, t_1]$, i.e., $\mathbf{P}(t_0 < T < t_1) = 1-\alpha$, where $\mathbf{P}(T<t_0) = \alpha/2$, and $\mathbf{P}(T>t_1) = \alpha/2$. These conditions give the following equations: $1 - e^{-t_0/600} = \alpha/2$, and $e^{-t_1/600} = \alpha/2$. Solving these equations, we have $t_0 = -600 \ln{(1 - \alpha/2)}$, and $t_1 = -600 \ln{(\alpha/2)}$.

For $\alpha = 10\%$, $t_0 = 30.77$ and $t_1 = 1797.44$, or $[0.51, 30.76]$ in minutes.

Let $T_1, T_2, T_3, \dots, T_n$ be the time to find the first block ($T_1$), then the time to find the second block ($T_2$), and so on. Let's analyze the distribution of $Y_n = \sum_{i=1}^{n} T_i \sim \text{Erlang}(n, \frac{1}{600})$. Thus, the CDF of $Y$ would be $\mathbf{P}(Y_n < t) = 1 - \sum_{k=0}^{n-1} \frac{1}{k!} e^{-\lambda t} (\lambda t)^k$.

Many exchanges use the rule of 6 confirmations. So, for $n=6$, we have: $\mathbf{P}(Y_6 < 1 \text{ hour}) = \mathbf{P}(Y_6 < 3600) = 0.5543$. The symmetrical confidence interval with level $\alpha=10\%$ is $[27, 105]$ in minutes.

% Probabilidade de orfão
An orphan block would be created if a new block is found during the propagation time of a new block. Let $\alpha$ be the percentage of the total hash rate of the node which is outdated, and $\Delta t$ the propagation time in seconds. Thus, $\mathbf{P}(\text{new orphan}) = \mathbf{P}(T < \Delta t) = 1 - e^{-\frac{\alpha \Delta t}{600}}$. For instance, if a node has 10\% of the total hash rate and it takes 30 seconds to receive the update, than $\mathbf{P}(\text{new orphan}) = 1 - e^{-\frac{0.1 \cdot 30}{600}} = 0.004987$.

\section{Analysis of network's hash rate change}
% Análise: Mudança no network's hash rate antes de atualizar o $A$.

The given number $A$ is adjusted every 2016 blocks. So, let's analyze what happens when the network's hash rate changes before the adjustment. First, $\mathbf{P}(13 \text{ days} < Y_{2016} < 15 \text{ days}) = \mathbf{P}(13 \cdot 24 \cdot 3600 < Y_{2016} < 14 \cdot 24 \cdot 3600) = 0.9986$, i.e., it will mostly happen between 13 and 15 days, assuming that the network's hash rate remains constant. But, as the network's total hash rate remained constant, the adjusted $A$ will not be changed.

%Let's assume that the network's total hash rate will change only once, from $H$ to $\alpha H$ -- so, if $\alpha = 1$, there would be no change, if $\alpha > 1$, there would be an increase, and if $\alpha < 1$, there would be a decrease. As the time required to find a new block would be $T_\alpha = \frac{X}{\alpha H}$, we got the same random variable as we have gotten when analyzing a miner with $\alpha$ percent of the network's total hash rate. But, the analysis will be different because here $\alpha$ may be greater than one.

Let $H \cdot u(t)$ be the network's total hash rate over time. So, the number of hashes calculated in $t$ seconds is $H \int_0^t u(t) dt$. Hence, $\mathbf{P}(T \leq t) = \mathbf{P}(X \leq H \int_0^t u(t)dt)$. When $H \rightarrow +\infty$, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{1}{600} \int_0^t u(t) dt}$, and the pdf of $T$ is $\frac{u(t)}{600} \cdot e^{-\frac{1}{600} \int_0^t u(t)dt}$.

Let's say that the network's total hash rate has suddenly multiplied by $\alpha$. So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ follows an exponential distribution with $\lambda = \frac{\alpha}{600}$. Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha$ follows an Erlang distribution. The time required to find 2016 blocks will be divided by $\alpha$, and the variance will be divided by $\alpha^2$. When the network's hash rate increases ($\alpha > 1$), the variance decreases. When the network's hash rate decreases ($\alpha < 1$), the variance increases.

For example, if $\alpha = 2$, $\mathbf{P}(6.5 \text{ days} < Y_{2016} < 7.5 \text{ days}) = 0.9986$, and the time required to find 2016 blocks halved. If $\alpha = 0.5$, $\mathbf{P}(27 \text{ days} < Y_{2016} < 29 \text{ days}) = 0.9469$, and the time required to find 2016 blocks doubled. So, even if half of the network stops mining, it will take at most 29 days to get back to normal.

% TODO Step-wise analysis

\subsection{Hash rate smoothly changing}

Let $u(t) = \frac{1+abx}{1+bx}$. It is an useful function because $u(0) = 1$ and $\lim_{t \rightarrow \infty} u(t) = a$. The bigger the $b$, the faster $u(t) \rightarrow a$. For example, if $a=2$, it means $H$ would be smoothly doubling. If $a=0.5$, it means $H$ would be smoothly halving.

It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a \Rightarrow \int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$

Assuming that $n = \frac{\lambda(a-1)}{b}$ is integer, we have:
$$F_T(t) = 1 - (1+bt)^n e^{-\lambda at}$$

Thus,
\begin{align*}
\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
\end{align*}

Hence, as $\mathcal{L}\{f_T(t)\} = s \mathcal{L}\{F_T(t)\}$,

$$\mathcal{L}\{f_T(t)\} = 1 - \sum_{k=0}^n \binom{n}{k} \frac{s b^k k!}{(s+\lambda a)^{k+1}}$$

Then,

\begin{align*}
\frac{d}{ds} \mathcal{L}\{f_T(t)\} 
	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{d}{ds} \frac{s}{(s+\lambda a)^{k+1}} \\
	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \left[ \frac{1}{(s+a\lambda)^{k+1}} - \frac{s(k+1)}{(s+a \lambda)^{k+1}} \right] \\
\frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0} 
	&= - \sum_{k=0}^n \binom{n}{k} b^k k! \frac{1}{(\lambda a)^{k+1}} \\
	&= - \frac{1}{a\lambda} \sum_{k=0}^n \binom{n}{k} k! \left( \frac{b}{\lambda a} \right)^k \\
	&= - \frac{1}{a\lambda} \sum_{k=0}^n \frac{n!}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \\
	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{(n-k)!} \left( \frac{b}{\lambda a} \right)^k \right] \\
	&= - \frac{1}{a\lambda} \left[ n! \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{n-k} \right] \tag{$k \rightarrow n-k$} \\
	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{b}{\lambda a} \right)^{-k} \right] \\
	&= - \frac{1}{a\lambda} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right]
\end{align*}

Finally, as $\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0}$,
$$\mathbf{E}[T] = \frac{1}{\lambda a} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right] \text{, where $n = \frac{\lambda(a-1)}{b}$}$$

Let's check this equation for already known scenarios. When $a=1$, then $n=0$ and $\mathbf{E}[T] = 1/\lambda$. When $b \rightarrow +\infty$, it reduces to the case in which the hash rate is multiplied by $a$, which we have already studied. In fact, $n \rightarrow 0$, $u(t) \rightarrow a$, and $\mathbf{E}[T] = \frac{1}{\lambda a}$.

\begin{theorem}
	$$a > 1, x > M \Rightarrow \left| \frac{1+abx}{1+bx} - a \right| < \frac{a-1}{1+bM}$$
\end{theorem}
\begin{proof}
$x > M \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM}$. As $1-a<0$, $\frac{1-a}{1+bx} > \frac{1-a}{1+bM}$. Thus, $\frac{1-a}{1+bM} < \frac{1-a}{1+bx} + a - a = \frac{1+abx}{1+bx} - a < 0 < \frac{a-1}{1+bM}$. Hence, $-\frac{a-1}{1+bM} < \frac{1+abx}{1+bx} - a < \frac{a-1}{1+bM}$.
\end{proof}

For instance, if we would like to know the impact of smoothly double the hash rate in the next week, $\lambda = 1/600$, $a=2$ and $M=1\text{ week}=3600\cdot24\cdot7 = 604,800$. $b$ can be calculated using $\epsilon = \frac{a-1}{1+bM} < 0.01$. So, $b > 0.000163690$ and $n<10.1818$. So, for $n=10$, $b=0.000166666$, and $\epsilon = 0.009823 < 0.01$, as expected. Finally, $\mathbf{E}[T] = 557.65$. If the hash rate had suddenly doubled, the average time between blocks would be 300.

%$$\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0} = \frac{b^n n! 600^{n+1}}{a^{n+1}} e^{\frac{a}{600b}} \text{, where $n = \frac{a-1}{600b}$}$$

%$a > 1, x > M \Rightarrow 1 + bx > 1 + bM \Rightarrow \frac{1}{1+bx} < \frac{1}{1+bM} \Rightarrow \frac{1-a}{1+bx} + a > \frac{1-a}{1+bM} + a \Rightarrow \frac{1+abx}{1+bx} - a > \frac{1-a}{1+bM}$.

\subsection{Hash rate changing linearly}

Let's analyze what would happen if the network's hash rate is growing linearly with angular coefficient $a^2$, i.e., $u(t) = 1 + a^2t$. Thus, $\mathbf{P}(T \leq t) = 1 - e^{-\frac{t + a^2t^2/2}{600}}$.

In order to obtain $\mathbf{E}(T)$, we will use the formula $\frac{d}{ds}T^*(s)|_{s=0} = -\mathbf{E}(T)$, where $T^*$ is the Laplace transform of the pdf of T.

Let $f_T(x)$ be the pdf of T and $F_T(x)$ be the cdf of T, i.e., $F_T(x) = \int_0^x f_T(t) dt$. Thus, $\mathcal{L}\{F_T(x)\} = \frac{T^*(s)}{s}$.

\begin{align*}
	\mathcal{L}\{F_T(x)\} &= \int_0^\infty e^{-sx} F_T(x) dx \\
		&= \int_0^\infty e^{-sx} \left( 1 - e^{-\frac{x + a^2x^2/2}{600}} \right) dx \\
		&= \int_0^\infty e^{-sx} dx - \int_0^\infty e^{-sx} e^{-\frac{x + a^2x^2/2}{600}} dx \\
		&= \frac{1}{s} - \int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{600}} dx \\
\end{align*}

Let $b=\frac{a}{20\sqrt{3}}$ and $c = \frac{10\sqrt{3}}{a} \left( s + \frac{1}{600} \right)$, then $sx + \frac{x+a^2x^2/2}{600} = (bx + c)^2 - c^2$.
Let $y = bx+c \Rightarrow dy = bdx$.

\begin{align*}
	\int_0^\infty e^{-sx -\frac{x + a^2x^2/2}{600}} dx &= \int_0^\infty e^{-(bx+c)^2 + c^2} dx \\
		&= \int_c^\infty e^{-y^2 + c^2} \frac{dy}{b} \\
		&= \frac{e^{c^2}}{b} \int_c^\infty e^{-y^2} dy \\
		&= \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c)
\end{align*}

$$ \mathcal{L}\{F_T(x)\} = \frac{1}{s} - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$

Hence,
$$ T^*(s) = s \mathcal{L}\{F_T(x)\} = 1 - \frac{s\sqrt{\pi}e^{c^2}}{2b} \erfc(c) $$

As $\frac{d}{ds}c = \frac{10\sqrt{3}}{a}$,

\begin{align*}
\frac{d}{ds} T^*(s) &= - \frac{\sqrt{\pi}e^{c^2}}{2b} \erfc(c) - \frac{20\sqrt{3\pi} sc e^{c^2}}{2ab} \erfc(c) + \frac{10\sqrt{3\pi}s}{2ab} \\
\frac{d}{ds} T^*(s)|_{s=0} &= - \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{1200a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right) \\
\mathbf{E}[T] &= \frac{10\sqrt{3\pi}}{a} e^{\frac{1}{1200a^2}} \erfc\left(\frac{\sqrt{3}}{60a}\right)
\end{align*}

It is easy to notice that $lim_{a \rightarrow 0} \mathbf{E}[T] = 600$, and $lim_{a \rightarrow \infty} \mathbf{E}[T] = 0$.

$\mathbf{E}[T]$ is ploted for $a \in (0, 0.1]$ at Figure \ref{fig-bitcoin-H-increase}. The plot interval is really small because $a=1$ would imply doubling the hash rate after 1 second.

\textcolor{red}{I do not know how to interpret this.}

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{images/bitcoin-H-increasing.png}
\caption{$\mathbf{E}[T]$ when H increases linearly with $u(t) = 1 + at$.\label{fig-bitcoin-H-increase}}
\end{figure}

\section{Attack in the Bitcoin network}
% Ataque submarino
% Probabilidade de "correr por fora"

Attackers are not part of the mining network, because they are trying to generate a different longest chain.

Let $\beta H$ be the hash rate of the attackers, and $\gamma H$ be the network's hash rate without the attackers. Thus, when $H \rightarrow +\infty$, we already know that $T_{\text{attackers}}$ and $T_{\text{network}}$ follow exponential distributions with parameters $\lambda_{\text{attacker}} = \frac{\beta}{600}$ and $\lambda_{\text{network}} = \frac{\gamma}{600}$, respectively.

As \cite{nakamoto2008bitcoin} have done, we will also model the attack using the Gambler's Ruin. In this game, a gambler wins \$1 at each round, with probability $p$, and loses \$1, with probability $1-p$. The rounds are independent. The gambler starts with \$$k$ plays continuously until he either accumulates a target amount of \$$m$, or loses all his money. Let $\rho = \frac{1-p}{p}$, then the probability of losing his fortune is:

$$
\mathbf{P}(\text{losing his fortune}) =
\begin{cases}
	\frac{\rho^k - \rho^m}{1-\rho^m} \text{, if $\rho \ne 1$,} \\
	\frac{m-k}{m} \text{, if $\rho = 1$.}
\end{cases}
$$

When $m \rightarrow +\infty$,

$$
\mathbf{P}(\text{losing his fortune}) =
\begin{cases}
	\rho^k \text{, if $\rho < 1$,} \\
	1 \text{, if $\rho \geq 1$.}
\end{cases}
$$

% TODO Expected time to absorption.

The gambler wins \$1 is the same as the network find a new block, the gambler losses \$1 is the same as the attacker find a new block. The initial \$$k$ is the same as the number of blocks the attacker is behind the network. Thus, the gambler losses his fortune is the same as the attacker successfully finds $k$ or more blocks than the network, i.e., losing his fortune means that the attack was successful.

In our case, $p = \frac{\lambda_{\text{network}}}{\lambda_{\text{network}} + \lambda_{\text{attacker}}} = \frac{\gamma}{\beta + \gamma}$, thus $\rho = \frac{\beta}{\gamma}$. Hence, $\rho < 1 \Leftrightarrow \beta < \gamma$.

Suppose that the attacker is mining with the network. Suddently, he stops mining with the network and starts attacking, i.e., starts to mine in another chain. In this scenario, since the attacker's hash rate is not mining with the network anymore, $\gamma = 1 - \beta$. Thus, $\beta < \gamma \Rightarrow \beta < 0.5 \Leftrightarrow \rho < 1$. Here comes the conclusion that, if the attacker has 50\% or more of the network's hash rate, then his attack will be certainly successful. We got exactly the same equations and conclusions as \cite{nakamoto2008bitcoin}.

But this scenario seems not to be the optimal attack, because the attacker has waited the $k$ confirmations before starting the attack. A better approach would be to start attacking just after propagating the transaction. In this case, our previous model is not good, because even if the attacker have found more blocks than the network, he cannot propagate those blocks before the network has found $k$ confirmations. So, we have to model the probabilities before the network has found the $k$ block. Then, if the attacker has more blocks than the network, he has successfully attacked. Otherwise, we got back to the previous model, in which the attacker must still find more blocks.

\begin{theorem}
	Let $a_{ij}$ be the probability of moving from state $(i, j)$ to state $(s, k)$ in the Markov chain.
	$$a_{ij} = \binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}.$$
\end{theorem}
\begin{proof}

In order to go from $(i, j)$ to $(s, k)$, the attacker must find exacly $i-s$ blocks, and the network must find exactly $k-j$ blocks. It is as they would be walking the grid from the point $(i, j)$ to $(s, k)$, where it is only allowed to go up or right, like in Figure \ref{figure:bitcoin-attack-paths}. There are many possible paths and all of them occurs with probability $(1-p)^{s-i} p^{k-j}$.

The walking ends when $(\cdot, k$) is reach, i.e., when the network finds $k$ blocks, no matter how many blocks the attacker has found. So, it is not allowed to walk through the line $(\cdot, k)$. Thus, the number of paths between $(i, j)$ and $(s, k)$ going only up or right, without going through the line $(\cdot, k)$ is exactly the number of paths between $(i, j)$ and $(s, k-1)$, which is equal to the number of permutations of the sequence $(u, u, \dots, u, r, r, \dots, r)$ in which there are $s-i$ right movements ($r$) and $k-j-1$ up movements ($u$). The number of permutations is $\frac{(k-j-1+s-i)!}{(s-i)!(k-j-1)!} = \binom{(k-j-1)+(s-i)}{s-i}$ because there are $s-i$ repetitions of the element $r$ and $k-j-1$ repetitions of the element $u$.

Finally, the probability is $\binom{k+s-1-i-j}{s-i} (1-p)^{s-i} p^{k-j}$.

\begin{figure}[ht]
  \centering
  \begin{tikzpicture}
    \coordinate (Origin)   at (0,0);
    \coordinate (XAxisMin) at (-1,0);
    \coordinate (XAxisMax) at (5,0);
    \coordinate (YAxisMin) at (0,-1);
    \coordinate (YAxisMax) at (0,5);
    %\draw [thin, gray,-latex] (XAxisMin) -- (XAxisMax);% Draw x axis
    %\draw [thin, gray,-latex] (YAxisMin) -- (YAxisMax);% Draw y axis

    \clip (-0.5,-0.5) rectangle (10.5cm,7cm); % Clips the picture...
    %\pgftransformcm{1}{0}{0}{1}{\pgfpoint{0cm}{0cm}}
          % This is actually the transformation matrix entries that
          % gives the slanted unit vectors. You might check it on
           % MATLAB etc. . I got it by guessing.
    \draw[style=help lines,dashed] (0,0) grid[step=1cm] (14,6);
          % Draws a grid in the new coordinates.
          %\filldraw[fill=gray, fill opacity=0.3, draw=black] (0,0) rectangle (2,2);
              % Puts the shaded rectangle
    \foreach \x in {0,1,...,10}{% Two indices running over each
      \foreach \y in {0,1,...,6}{% node on the grid we have drawn 
        \node[draw,circle,inner sep=1pt,fill] at (\x,\y) {};
            % Places a dot at those points
      }
    }

    \draw [ultra thick,-latex,red] (0,0) -- (0,1) node [midway, left] {};
    \draw [ultra thick,-latex,red] (0,1) -- (0,2) node [midway, left] {};
    \draw [ultra thick,-latex,red] (0,2) -- (1,2) node [midway, above] {};
    \draw [ultra thick,-latex,red] (1,2) -- (1,3) node [midway, right] {};
    \draw [ultra thick,-latex,red] (1,3) -- (2,3) node [midway, right] {};
    \draw [ultra thick,-latex,red] (2,3) -- (2,4) node [midway, right] {};
    \draw [ultra thick,-latex,red] (2,4) -- (2,5) node [midway, right] {};
    \draw [ultra thick,-latex,red] (2,5) -- (3,5) node [midway, right] {};
	\draw [ultra thick,-latex,red] (3,5) -- (3,6) node [above] {(3, 6)};

    \draw [ultra thick,-latex,blue] (0,0) -- (1,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (1,0) -- (2,0) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,0) -- (2,1) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,1) -- (2,2) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,2) -- (2,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (2,3) -- (3,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (3,3) -- (4,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (4,3) -- (5,3) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,3) -- (5,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (5,4) -- (6,4) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,4) -- (6,5) node [midway, left] {};
    \draw [ultra thick,-latex,blue] (6,5) -- (7,5) node [midway, left] {};
	\draw [ultra thick,-latex,blue] (7,5) -- (7,6) node [above] {(7,6)};
  \end{tikzpicture}
  \caption{Both the attacker and the network are mining. Each step up is a new block found by the network with probability $p$. Each step right is a new block found by the attacker with probability $1-p$. It ends when the network finds $k$ blocks -- in this example, $k=6$. The red path has probability $p^6 (1-p)^3$, while the blue path has probability $p^6 (1-p)^7$. Notice that the blue path is a successfull attack, because the attacker has found more blocks than the network. In the red path, the attacker still have to catch up 3 blocks to have a successfull attack, which happens with probability $\rho^3$, if $p < 0.5$.}
  \label{figure:bitcoin-attack-paths}
\end{figure}
\end{proof}

\begin{cor}
	Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found exactly $s$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S = s) = \binom{k+s-1}{s} (1-p)^s p^k.$
\end{cor}
\begin{proof}
	This probability $\mathbf{P}(S = s) = a_{00}$.
\end{proof}

Assuming that the attacker starts the attack just after publishing the transaction, the probability of the attacker has already found more than $k$ blocks while it waits the network to find $k$ blocks is $\mathbf{P}(S \geq k) = \sum_{s=k}^{\infty} \binom{k+s-1}{s} (1-p)^s p^k$.
